---
title: "Writing Reports with R Markdown"
author: "Paige"
date: "2022-10-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, include=FALSE}
library(tidyverse)
library(here)
library(knitr)
```

This report was prepared for the UN. It analyzes the relationship between a country's GDP, life expectancy, and CO2 emissions. Our goal was to determine to what degree a country's economic strength or weakness may be related to its public health status and impact on climate pollution. We hypothesize that both life expectancy and CO2 emissions will increase with a country's GDP.

```{r data, include=FALSE}
gapminder_1997 <- read_csv(here("data/gapminder_1997.csv"))
```

```{r gdp_lifeExp1997, echo=FALSE}
ggplot(gapminder_1997, aes(x=gdpPercap, y=lifeExp, color=continent, size=pop/1000000))+
  labs(x="GDP Per Capita", y="Life Expectancy", title = "Do People in Wealthy Countries Live Longer?", size= "Population in Millions")+
  geom_point()
```

``` {r summary data, echo=FALSE}
n_countries <- gapminder_1997%>%
  select(country)%>%
  n_distinct()
n_minGDP <- min(gapminder_1997$gdpPercap)%>%round(digits = 2)
n_maxGDP <- max(gapminder_1997$gdpPercap)%>%round(digits = 2)
```


The above plot shows the relationship between GDP per capita and life expectancy for a total of `r n_countries` countries. For this set of countries, economic wealth ranged from a minimum of $`r n_minGDP` to a maximum of $`r n_maxGDP`. 


|Summary of Data |
|------------|---------|
|Number of Countries | `r n_countries`|
|Minimum GDP per Capita| `r n_minGDP`|
|Maximum GDP per Capita | `r n_maxGDP`|

```{r tables}
gapminder_1997%>%
  filter(continent == "Oceania")%>%
  kable()

```

### Lists unordered
1. fave language is R because idk any others
2. *i need to re learn stats but so far the tidyr cleaning columns and mutate or select is good* 
3. ***i would love to be good at GIS i've been getting better but it is hard to work with census data and tables in GIS***
4. [r markdown cheatsheet](chrome-extension://efaidnbmnnnibpcajpcglclefindmkaj/https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf)

### ordered list
- dsfsfs
    - djff
    - kjkjsdf